Simulation study in rectangular domain

### 2024-04-27 R/simulation_study/simulation_study_fjords.Rmd
changes in simulation study code

```{r import libraries}
#seed for reproducibility 
set.seed(1)

@@

```{r set observation error}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs=0.005
H_hf=array(rep(sigma_obs^2*diag(2),n_hf),dim=c(2,2,n_hf))

@@

ftau_bump=function(cov_data,epsilon=pi/10,alpha=pi/5,D0=1) {
    Dshore=cov_data$DistanceShore
    theta=cov_data$theta
    tau=(1.5+0.5*(tanh((theta+(pi/2+epsilon))/alpha)-tanh((theta+(pi/2-epsilon))/alpha))+
      0.5*(tanh((theta+(-pi/2+epsilon))/alpha)-tanh((theta+(-pi/2-epsilon))/alpha)))
    return (tau)
}

fomega=function(cov_data,D0=0.1,omega0=60*pi,lambda=2,kappa=0.2) {
    Dshore=cov_data$DistanceShore
    theta=cov_data$theta
    if (is.null(Dshore)){
      Dshore=1/cov_data$ExpShore
    }
    coeff=exp(-kappa*(Dshore/D0)^2)
    omega=omega0/2*(tanh(lambda*(theta+pi/2-atanh(-0.9)/lambda))+tanh(lambda*(theta-(pi/2-atanh(-0.9)/lambda))))*coeff
    return(omega)
}

@@

Dshore=seq(from=0.1,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")
true_tau=matrix(ftau_constant(grid),30,30)
true_omega=matrix(fomega(grid),30,30)

@@

@@

#sample theta and DistanceShore points in the domain
theta <- runif(n,-pi,pi);
DistanceShore <- runif(n,0.2,2);
samples=data.frame(theta=theta,DistanceShore=DistanceShore)

#define grid of values
theta_v <- seq(-pi,pi,length=30)
Dshore_v<- seq(0.2,2,length=30)
pr <- data.frame(theta=rep(theta_v,30),DistanceShore=rep(Dshore_v,rep(30,30)))

@@

plot_spline_omega <- plot_ly(type = "surface",
                             contours = list(z = list(show = TRUE, start = min(spline_omega), end = max(spline_omega), size = 5,color="black")),
                         x = ~Dshore,y = ~theta,z=~spline_omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline smooth omega"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis2 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z2), end = max(z2), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z2, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 2"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis3 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z3), end = max(z3), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z3, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 3"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis4 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z4), end = max(z4), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z4, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 4"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega"))) 
```

@@

@@

ggsave(filename="samples_fjords_hf_ne_omega.png",plot=plot_samples_fjords_hf_ne)
ggplotly(plot_samples_fjords_hf_ne)
```

>>>>

```{r import libraries}
#seed for reproducibility 
set.seed(42)

@@

```{r set observation error}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs=0.01
H_hf=array(rep(sigma_obs^2*diag(2),n_hf),dim=c(2,2,n_hf))

@@

ftau_bump=function(cov_data,epsilon=pi/10,alpha=pi/5,D0=1,tau0=0.8,tau1=2) {
    Dshore=cov_data$DistanceShore
    theta=cov_data$theta
    tau=(tau0+(tau1-tau0)*(tanh((theta+(pi/2+epsilon))/alpha)-tanh((theta+(pi/2-epsilon))/alpha))+
      (tau1-tau0)*(tanh((theta+(-pi/2+epsilon))/alpha)-tanh((theta+(-pi/2-epsilon))/alpha)))
    return (tau)
}

fomega=function(cov_data,D0=0.3,omega0=60*pi/2,lambda=2,kappa=0.2) {
    Dshore=cov_data$DistanceShore
    theta=cov_data$theta
    if (is.null(Dshore)){
      Dshore=1/cov_data$ExpShore
    }
    coeff=exp(-kappa*(Dshore/D0)^2)
    omega=omega0/2*(tanh(lambda*(theta+pi/2-atanh(-0.9)/lambda))+tanh(lambda*(theta-(pi/2-atanh(-0.9)/lambda))))*coeff
    return(omega)
}

@@

Dshore=seq(from=0.1,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")
true_tau=matrix(ftau_bump(grid),30,30)
true_omega=matrix(fomega(grid),30,30)

@@

saveWidget(plot_true_omega, file ="true_smooth_omega_fjords.html")

@@

#sample theta and DistanceShore points in the domain
theta <- runif(n,-pi,pi);
DistanceShore <- runif(n,0.2,3);
samples=data.frame(theta=theta,DistanceShore=DistanceShore)

#define grid of values
theta_v <- seq(-pi,pi,length=30)
Dshore_v<- seq(0.2,3,length=30)
pr <- data.frame(theta=rep(theta_v,30),DistanceShore=rep(Dshore_v,rep(30,30)))

@@

plot_spline_omega <- plot_ly(type = "surface",
                             contours = list(z = list(show = TRUE, start = min(spline_omega), end = max(spline_omega), size = 5,color="black")),
                         x = ~theta,y = ~Dshore,z=~spline_omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline smooth omega"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis2 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z2), end = max(z2), size = 0.05,color="black")),
                         x = ~theta,y = ~Dshore,z=~z2, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 2"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis3 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z3), end = max(z3), size = 0.05,color="black")),
                         x = ~theta,y = ~Dshore,z=~z3, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 3"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis4 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z4), end = max(z4), size = 0.05,color="black")),
                         x = ~theta,y = ~Dshore,z=~z4, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 4"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

saveWidget(plot_spline_omega, file ="spline_smooth_omega_fjords.html")
```

@@

```{r}
res=sim_theta_CRCVM(ftau=ftau_bump,fomega=fomega,fnu=fnu_constant,v0=v0,x0=x0[2,],times=seq(0,5*24,by=1/60),land=land,verbose=FALSE)

data_sim=res$sim
data_sim$ID=factor(rep(1,length(data_sim$y1)))
data_shore=res$shore[,c("p1","p2")]
data_sim=cbind(data_sim,data_shore)

data_sim=data_sim[seq(1,length(data_sim$time),by=5),]

plot_illust_fjords=ggplot()+geom_sf(data=land$geometry,fill="grey")+coord_sf(datum=st_crs(32626))+
   geom_point(data=data_sim,aes(p1,p2),shape=5,size=0.1,col="red")+
  geom_point(data=data_sim,mapping=aes(y1,y2),size=0.1)+
  geom_path(data=data_sim,mapping=aes(y1,y2),size=0.1)+
  geom_point(data = data_sim%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

ggplotly(plot_illust_fjords)

ggsave(filename="illustrative_sample_fjords.png",plot=plot_illust_fjords,width=10,height=5)
```

@@

ggsave(filename="samples_fjords_hf_ne_omega.png",plot=plot_samples_fjords_hf_ne,width=10,height=5)
ggplotly(plot_samples_fjords_hf_ne)
```

### 2024-04-27 R/simulation_study/simulation_study_rect.Rmd
changes in simulation study code

```{r measurement error}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs=0.005
H_hf=array(rep(sigma_obs^2*diag(2),n_hf*N_ID),dim=c(2,2,n_hf*N_ID))

@@

plot_true_tau <- plot_ly(type = "surface",
                         x = ~Dshore,y = ~theta,z=~true_tau,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True tau"),
                       scene=list(xaxis = list(title = "Theta",showgrid = F),
                      yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "tau")))

plot_true_omega <- plot_ly(type = "surface",
                           contours = list(z = list(show = TRUE, start = min(true_omega), end = max(true_omega), size = 5,color="black")),
                         x = ~Dshore,y = ~theta,z=~true_omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True omega"),
                       scene=list(xaxis = list(title = "Theta",showgrid = F),
                      yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

plot_true_tau
plot_true_omega
```

@@

plot_spline_omega <- plot_ly(type = "surface",
                             contours = list(z = list(show = TRUE, start = min(spline_omega), end = max(spline_omega), size = 5,color="black")),
                         x = ~Dshore,y = ~theta,z=~spline_omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline smooth omega"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis2 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z2), end = max(z2), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z2, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 2"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis3 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z3), end = max(z3), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z3, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 3"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis4 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z4), end = max(z4), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z4, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 4"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega"))) 
```

@@

```{r}
plot_basis2
```

```{r}
plot_basis3
```

```{r}
plot_basis4
```

@@

### Generate sample

```{r parallel generate hf low errors samples,cache=TRUE}

@@

### plot one batch

@@

plot_samples_rect_hf_ne=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_path(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_point(data = data_rect_hf_ne$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

ggsave(filename="samples_rect_hf_ne_omega.png",plot=plot_samples_rect_hf_ne)
ggplotly(plot_samples_rect_hf_ne)
```

>>>>

```{r measurement error}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs=0.01
H_hf=array(rep(sigma_obs^2*diag(2),n_hf*N_ID),dim=c(2,2,n_hf*N_ID))

@@

plot_true_tau <- plot_ly(type = "surface",
                         x = ~theta,y = ~Dshore,z=~true_tau,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True tau"),
                       scene=list(xaxis = list(title = "Theta",showgrid = F),
                      yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "tau")))

plot_true_omega <- plot_ly(type = "surface",
                           contours = list(z = list(show = TRUE, start = min(true_omega), end = max(true_omega), size = 5,color="black")),
                         x = ~theta,y = ~Dshore,z=~true_omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True omega"),
                       scene=list(xaxis = list(title = "Theta",showgrid = F),
                      yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

plot_true_tau
plot_true_omega

saveWidget(plot_true_omega, file ="true_smooth_omega_rect.html")

@@

plot_spline_omega <- plot_ly(type = "surface",
                             contours = list(z = list(show = TRUE, start = min(spline_omega), end = max(spline_omega), size = 5,color="black")),
                         x = ~theta,y = ~Dshore,z=~spline_omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline smooth omega"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis2 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z2), end = max(z2), size = 0.05,color="black")),
                         x = ~theta,y = ~Dshore,z=~z2, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 2"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis3 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z3), end = max(z3), size = 0.05,color="black")),
                         x = ~theta,y = ~Dshore,z=~z3, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 3"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis4 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z4), end = max(z4), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z4, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 4"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

saveWidget(plot_spline_omega, file ="spline_smooth_omega_rect.html")
```

@@

```{r}
plot_basis2

saveWidget(plot_basis2, file ="spline_basis2_omega_rect.html")
```

```{r}
plot_basis3
saveWidget(plot_basis3, file ="spline_basis3_omega_rect.html")
```

```{r}
plot_basis4
saveWidget(plot_basis4, file ="spline_basis4_omega_rect.html")
```

@@

### Generate samples

```{r illustrative sample for the paper}

res=sim_theta_CRCVM(ftau=ftau_constant,fomega=fomega,fnu=fnu_constant,v0=v0,x0=x0[1,],times=seq(0,2*24,by=1/60),land=domain,verbose=FALSE)

data_sim=res$sim
data_sim$ID=factor(rep(1,length(data_sim$y1)))
data_shore=res$shore[,c("p1","p2")]
data_sim=cbind(data_sim,data_shore)

data_sim=data_sim[seq(1,length(data_sim$time),by=5),]

plot_illust_rect=ggplot()+geom_sf(data=domain$geometry,fill="grey")+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_sim,aes(p1,p2),shape=5,size=0.1,col="red")+
  geom_point(data=data_sim,mapping=aes(y1,y2),size=0.1)+
  geom_path(data=data_sim,mapping=aes(y1,y2),size=0.1)+
  geom_point(data = data_sim%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

plot_illust_rect

ggsave(filename="illustrative_sample_rect.png",plot=plot_illust_rect,width=10,height=5)
```

```{r parallel generate batches of hf low errors samples,cache=TRUE}

@@

### plot one sample and one batch

@@

plot_samples_rect_hf_ne=ggplot()+geom_sf(data=domain$geometry,fill="grey")+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID,shape=ID),size=0.1)+
  geom_path(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID,shape=ID),size=0.1)+
  geom_point(data = data_rect_hf_ne$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

ggsave(filename="samples_rect_hf_ne_omega.png",plot=plot_samples_rect_hf_ne,width=10,height=5)
ggplotly(plot_samples_rect_hf_ne)
```
